\selectlanguage{english}%
\noindent \begin{center}
	\textbf{\Large{}RDFplay: a playground for education on the Semantic Web technologies}
	\par\end{center}{\Large \par}

\noindent \begin{center}
\textbf{\Large{}Abstract}
\par\end{center}{\Large \par}


The aim of this project is to develop a playground system for \gls{SPARQL} and \gls{RDF} that will serve as an educational tool for introducing Semantic Web Technologies. Its purpose is to provide a user-friendly interface for profane users to have their first contact in this kind of technologies at a beginner level, allowing them to carry out academic activities on simple graph manipulation or \gls{SPARQL} queries to the work dataset and to external endpoints.

While it is true that the Semantic Web today reaches a reasonably advanced stage of maturity, the rate of change of frontend technologies in \textit{Javascript} (or JS) is, at least, vertiginous. Moreover, most of the Semantic Web components implementations have been developed in backend technologies such as \textit{Java} (\textit{Apache Jena}) or \textit{Python} (\textit{rdflib}), with a lack of existing standards or pure \textit{Javascript} mature implementations yet.

Three challenges are therefore addressed: the need to obtain a product both simple and easy to use by non-expert users, the difficulty of finding mature components combining the Semantic Web and frontend and the convenience of betting on a stable \textit{Javascript} development framework with a smooth learning curve.

To solve these problems, the choice made is to develop a \gls{SPA} with \textit{Vue.js} (a JS framework that boasts of agglutinating the best features of \textit {Angular} and \textit {React}, its main competitors in the market) and integrate it with the \textit{RDF Javascript libraries working group} recommended implementations and a flexible and modular web query platform. 

Given the high uncertainty found when implementing the project needs, an incremental and iterative methodological approach based on \textit {Extreme Programming} and supported on Kanban boards has been used, which has led to a better project organization and evolution.

The final result provides for a Semantic Web technologies introduction basic product and allows to think of a development as ambitious as the own teaching teams needs, given the current maturity state of the web \textit {frontend}.

\textbf{Keywords:} education, \textit{Javascript}, Semantic Web, \textit{Vue.js}, \gls{RDF}, \gls{SPARQL}, ontologies.

\selectlanguage{spanish}%

